---
title: "clustering_time_series"
author: "KiseokUchicago"
date: "2021-03-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

### Time series clustering

### Unsupervised clustering with various distances (SBD, DTW, ...) / prototyping methods / clustering methods \
- Using dtwclust package
- How many clusters should we use? Evaluate by CVI (Cluster validity index, the bigger the better for internal CVI) \
- Please normalize before clustering

### 1. Data formating for tsclust input 

```{r}
# libraries
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(formattable)
library(dtwclust)
require("TSclust")
```

Input data formating

```{r}
# read in combined table of 24 samples (Replicate B)
df_pep <- read.table("Replicate_ABC67_combined.tsv",sep='\t',header = T)

# Let's try to plot all of them at once
df_all <- df_pep %>% select(Peptide_id, SampleID, Time, Depth) %>% unique() %>% arrange(Peptide_id)
# How many time points did it appear
(df_all2 <- df_all %>% group_by(Peptide_id, Depth) %>% summarize(Timepoints = n_distinct(Time)) %>% ungroup())
df_all2$Depth <- factor(df_all2$Depth, levels = c('SRF','DCL','BDCL'))

# (1) Get list of peptides that are present in 7 or 8 timepoints
df_all2_filt <- df_all2 %>% filter(Timepoints > 6) %>% unique() 

# (2) Let's average the abundance_ratio for each unique peptide
df_int <- df_pep %>% select(Peptide_id, Spec_id, cent.log2.16O.18O.Ratio, SampleID, Time, Depth, Replicate) %>%
  unique() %>% arrange(Peptide_id)
# first average the spectrum_id per sample 
# important to make sure to take the mean of the log raio (instead of taking log2 after averaging the original value(2^x))
df_ave <- df_int %>% group_by(SampleID, Time, Depth, Replicate, Peptide_id) %>%
  summarize(Mean_abundance_ratio=mean(cent.log2.16O.18O.Ratio),
            SD_pep=sd(cent.log2.16O.18O.Ratio)) %>% ungroup()
head(df_ave)
dim(df_ave)
# Now, average and error by replicate
# see here for calculating error bar in log scale (https://faculty.washington.edu/stuve/log_error.pdf)
# but here we will not use the method in the link above
# calculate average and error bar for abundance_ratio
# SE calculated with Jake Waldbauer's method: sqrt(se(Mean_abundance_ratio)^2 +(mean(SD_pep))^2))
# standard error
se <- function(x) sd(x)/sqrt(length(x))
df_ave_rep <- df_ave %>% group_by(Time, Depth, Peptide_id) %>%
  summarize(Ave_abundance_ratio=mean(Mean_abundance_ratio),
            SE_abundance_ratio=sqrt((se(Mean_abundance_ratio))^2 +(mean(SD_pep))^2)) %>% ungroup()
head(df_ave_rep,10)
# why is there no sd for Pep_1004?
df_ave %>% group_by(SampleID) %>% filter(Peptide_id=='Pep_1003') # there is only 1 sample for T1 BDCL 

df_ave_rep2 <- df_ave_rep %>% mutate(error_up=Ave_abundance_ratio+SE_abundance_ratio) %>%
  mutate(error_down=Ave_abundance_ratio-SE_abundance_ratio)

head(df_ave_rep2)

# (3) import time metadata and merge
(df_time <- read.table('data/Time_metadata.txt', header=T, sep='\t'))
df_time$date_time <- str_replace(df_time$date_time, "/2019"," ")
df_time$date_time <- str_replace(df_time$date_time, ":00$","")

# left_join with average data
df_ave2 <- df_ave_rep2 %>% left_join(df_time, by=c('Time'='timepoint'))
head(df_ave2)
# only use peptides that appear 7 or 8 times
df_ave2 %>% select(Peptide_id) %>% unique() %>% dim() # 3233 peptides before
df_filt78 <- inner_join(df_ave2, df_all2_filt, by=c('Peptide_id'='Peptide_id','Depth'='Depth'))

# inspecting if the inner join was successful
head(df_filt78)
colnames(df_ave2)
colnames(df_time)
colnames(df_filt78)

df_filt78 %>% select(Peptide_id) %>% unique() %>% dim() # 841 peptides for peptides in 7 or 8 timepoints
df_filt78$Depth <- factor(df_filt78$Depth, levels = c('SRF','DCL','BDCL'))

# get only 8 ones
df_filt8 <- df_filt78 %>% filter(Timepoints > 7)
df_filt8 %>% select(Peptide_id) %>% unique() %>% dim() # 661 peptides for peptides in all 8 timepoints
df_filt8$Depth <- factor(df_filt8$Depth, levels = c('SRF','DCL','BDCL'))

# get only 7 ones
df_filt7 <- df_filt78 %>% filter(Timepoints == 7)
df_filt7 %>% select(Peptide_id) %>% unique() %>% dim() # 356 peptides for peptides in 7 timepoints
df_filt7$Depth <- factor(df_filt7$Depth, levels = c('SRF','DCL','BDCL'))

```


```{r}
# (1) Change our data into list format for input into tsclust
colnames(df_filt8)

make_list <- function(df_filt8,depth){
  df_SRF8 <- df_filt8 %>% filter(Depth==depth) %>% select(Peptide_id,Time,Ave_abundance_ratio)
  # reshape the dataframe with timepoints
  library(maditr)
  v_names <- c("T1","T2","T3","T4","T5","T6","T7","T8")
  df_SRF8$Time <- factor(df_SRF8$Time, levels=v_names)
  dc_SRF8 <- dcast(df_SRF8, Peptide_id ~ Time, value.var = "Ave_abundance_ratio")
  # make Peptide_id more simpler (Pep_1 -> P.1)
  dc_SRF8$Peptide_id <- str_replace(dc_SRF8$Peptide_id,"Pep_","P.")
  # make first column into row name
  dc_SRF8 <- tibble::column_to_rownames(dc_SRF8, var="Peptide_id")
  # make it into list format
  # l_SRF8 <- setNames(split(dc_SRF8, seq(nrow(dc_SRF8))), rownames(dc_SRF8))
  l_SRF8 <- as.list(as.data.frame(t(dc_SRF8)))
  return(l_SRF8)
}

l_SRF8 <- make_list(df_filt8,'SRF')
l_DCL8 <- make_list(df_filt8,'DCL')
l_BDCL8 <- make_list(df_filt8,'BDCL')

l_SRF8_z <- zscore(l_SRF8)
l_DCL8_z <- zscore(l_DCL8)
l_BDCL8_z <- zscore(l_BDCL8)


```

### 2. Hierarchical clustering w/ shape based distance

```{r}
# input into tsclust
library(dtwclust)
require("TSclust")

### Hierarchical clustering w/ shape based distance
hc_sbd <- tsclust(l_SRF8, type = "h", k = 2L:10L,
  preproc = zscore, seed = 899,
  distance = "sbd", centroid = shape_extraction,
  control = hierarchical_control(method = "average"))

internal_CVI <- function(hc_sbd,main){
  names(hc_sbd) <- 2L:10L
  (eval <- sapply(hc_sbd, cvi, type = "internal"))
  # divide CH value with 100 just to plot with other internal CVIs 
  eval['CH',] <- eval['CH',]/100
  t(eval)
  matplot(x=seq(2,10,by=1),y=t(eval), type = c("b"),pch=1,col = 1:7,
          xlab="Number of clusters",ylab="Internal CVI (cluster validity index)",main=main,axes=F) #plot
  axis(2)
  axis(side=1,at=2:10)
  legend("bottomright", legend = c("Sil","SF","CH/100","DB","DBstar","D","COP"), col=1:7, pch=1,cex=0.8) # optional legend
}

internal_CVI(hc_sbd,"Hierarchical clustering w/ shape based distance")

# use 4 or 5 or 6 clusters
hc_sbd4 <- tsclust(l_SRF8, type = "h", k = 4L,
  distance = "sbd", centroid = shape_extraction,
  control = hierarchical_control(method = "average"))

hc_sbd5 <- tsclust(l_SRF8, type = "h", k = 5L,
  distance = "sbd", centroid = shape_extraction,
  control = hierarchical_control(method = "average"))

hc_sbd6 <- tsclust(l_SRF8, type = "h", k = 6L,
  distance = "sbd", centroid = shape_extraction,
  control = hierarchical_control(method = "average"))

# cluster dendrogram
plot(hc_sbd4)
# The series and the obtained prototypes can be plotted too
plot(hc_sbd4, type = "sc")
# Focusing on the first cluster
plot(hc_sbd4, type = "series", clus = 1L)
plot(hc_sbd4, type = "centroids", clus = 1L)

# cluster dendrogram
plot(hc_sbd5)
# The series and the obtained prototypes can be plotted too
plot(hc_sbd5, type = "sc")

# cluster dendrogram
plot(hc_sbd6)
# The series and the obtained prototypes can be plotted too
plot(hc_sbd6, type = "sc")
```

### 3.1. partial clustering methods - DTW

```{r}
### partial clustering methods
# z-normalization
l_SRF8_z <- zscore(l_SRF8)

# DTW distance
pc_dtw <- tsclust(l_SRF8_z, k = 2L:10L,
  distance = "dtw_basic", centroid = "dba",
  trace = TRUE, seed = 8,
  norm = "L2", window.size = 8L,
  args = tsclust_args(cent = list(trace = TRUE)))

internal_CVI(pc_dtw, "Partial clustering with DTW distance")

# use 2, 3, 4 clusters

pc_dtw2 <- tsclust(l_SRF8_z, k = 2L,
  distance = "dtw_basic", centroid = "dba",
  trace = TRUE, seed = 8,
  norm = "L2", window.size = 8L,
  args = tsclust_args(cent = list(trace = TRUE)))
pc_dtw3 <- tsclust(l_SRF8_z, k = 3L,
  distance = "dtw_basic", centroid = "dba",
  trace = TRUE, seed = 8,
  norm = "L2", window.size = 8L,
  args = tsclust_args(cent = list(trace = TRUE)))
pc_dtw4 <- tsclust(l_SRF8_z, k = 4L,
  distance = "dtw_basic", centroid = "dba",
  trace = TRUE, seed = 8,
  norm = "L2", window.size = 8L,
  args = tsclust_args(cent = list(trace = TRUE)))



# Prototypes for 2 clusters
plot(pc_dtw2)
plot(pc_dtw2, type = "centroids")

# Prototypes for 3 clusters
plot(pc_dtw3)
plot(pc_dtw3, type = "centroids")

# Prototypes for 4 clusters
plot(pc_dtw4)
plot(pc_dtw4, type = "centroids")

```

### 3.2. partial clustering methods - DTW_lb
cross-distance matrix is NOT symmetric, which can be problematic for

```{r}
# DTW_lb distance method
pc_dtwlb <- tsclust(l_SRF8_z, k = 2L:10L,
  distance = "dtw_lb", centroid = "dba",
  trace = TRUE, seed = 8,
  norm = "L2", window.size = 4L,
  control = partitional_control(pam.precompute = FALSE),
  args = tsclust_args(cent = list(trace = TRUE)))

internal_CVI(pc_dtwlb, "Partial clustering with DTW_lb distance")

# use 8 clusters
pc_dtwlb8 <- tsclust(l_SRF8_z, k = 8L,
  distance = "dtw_lb", centroid = "dba",
  trace = TRUE, seed = 8,
  norm = "L2", window.size = 4L,
  control = partitional_control(pam.precompute = FALSE),
  args = tsclust_args(cent = list(trace = TRUE)))

# Prototypes for 8 clusters
plot(pc_dtwlb8)
plot(pc_dtwlb8, type = "centroids")

# use 4 clusters
pc_dtwlb4 <- tsclust(l_SRF8_z, k = 4L,
  distance = "dtw_lb", centroid = "dba",
  trace = TRUE, seed = 8,
  norm = "L2", window.size = 4L,
  control = partitional_control(pam.precompute = FALSE),
  args = tsclust_args(cent = list(trace = TRUE)))

# Prototypes for 8 clusters
plot(pc_dtwlb4)
plot(pc_dtwlb4, type = "centroids")


```

### 3.3. partial clustering methods - k-shape

```{r}
# k-shape method
pc_ks <- tsclust(l_SRF8_z, k = 2L:10L,
  distance = "sbd", centroid = "shape",
  seed = 8, trace = TRUE)

internal_CVI(pc_ks, "Partial clustering with k-shape method")

# use 2, 3, or 4 clusters
pc_ks2 <- tsclust(l_SRF8_z, k = 2L,
  distance = "sbd", centroid = "shape",
  seed = 8, trace = TRUE)

pc_ks3 <- tsclust(l_SRF8_z, k = 3L,
  distance = "sbd", centroid = "shape",
  seed = 8, trace = TRUE)

pc_ks4 <- tsclust(l_SRF8_z, k = 4L,
  distance = "sbd", centroid = "shape",
  seed = 8, trace = TRUE)

pc_ks5 <- tsclust(l_SRF8_z, k = 5L,
  distance = "sbd", centroid = "shape",
  seed = 8, trace = TRUE)

# Prototypes for 2 clusters
plot(pc_ks2)
plot(pc_ks2, type='centroid')

# Prototypes for 3 clusters
plot(pc_ks3)
plot(pc_ks3, type='centroid')

# Prototypes for 4 clusters
plot(pc_ks4)
plot(pc_ks4, type='centroid')

# Prototypes for 5 clusters
plot(pc_ks5)
plot(pc_ks5, type='centroid')

```


### 3.4. partial clustering methods - Tadpole 
problematic for our dataset (series' cross-distance matrix is NOT symmetric)

```{r}
# Tadpole method
pc_tp <- tsclust(l_SRF8_z, k = 2L:10L, type = "t",
  seed = 8, trace = TRUE,
  control = tadpole_control(dc = 1.5,
  window.size = 4L))

internal_CVI(pc_tp, "Partial clustering with Tadpole method")

# use 4 or 5 clusters
pc_tp4 <- tsclust(l_SRF8_z, k = 4L, type = "t",
  seed = 8, trace = TRUE,
  control = tadpole_control(dc = 1.5,
  window.size = 4L))

pc_tp5 <- tsclust(l_SRF8_z, k = 5L, type = "t",
  seed = 8, trace = TRUE,
  control = tadpole_control(dc = 1.5,
  window.size = 4L))

# Prototypes for 4,5 clusters
plot(pc_tp4)
plot(pc_tp4, type="centroid")

plot(pc_tp5)
plot(pc_tp5, type="centroid")

```

### 4. Using Tadpole for other depths

(2) DCL

```{r}
# DCL
pc_tp <- tsclust(l_DCL8_z, k = 2L:10L, type = "t",
  seed = 8, trace = TRUE,
  control = tadpole_control(dc = 1.5,
  window.size = 4L))

internal_CVI(pc_tp, "Partial clustering with Tadpole method")

# use 4 or 5 clusters
partial_tadpole <- function(df_filt8, depth="DCL",k=4){
  l_DCL8 <- make_list(df_filt8,'DCL')
  l_DCL8_z <- zscore(l_DCL8)
  pc_tp_k <- tsclust(l_DCL8_z, k = k, type = "t",
    seed = 8, trace = TRUE,
    control = tadpole_control(dc = 1.5,
    window.size = 4L))
  plot(pc_tp_k)
  plot(pc_tp_k, type="centroid")
}

# Prototypes for DCL
partial_tadpole(df_filt8, depth="DCL",k=3)
partial_tadpole(df_filt8, depth="DCL",k=4)
partial_tadpole(df_filt8, depth="DCL",k=5)
partial_tadpole(df_filt8, depth="DCL",k=6)

```

(3) BDCL

```{r}
# BDCL
pc_tp <- tsclust(l_BDCL8_z, k = 2L:10L, type = "t",
  seed = 8, trace = TRUE,
  control = tadpole_control(dc = 1.5,
  window.size = 4L))

internal_CVI(pc_tp, "Partial clustering with Tadpole method")

# Prototypes for BDCL
partial_tadpole(df_filt8, depth="BDCL",k=2)
partial_tadpole(df_filt8, depth="BDCL",k=3)
partial_tadpole(df_filt8, depth="BDCL",k=4)
partial_tadpole(df_filt8, depth="BDCL",k=5)
partial_tadpole(df_filt8, depth="BDCL",k=6)

```

### 4.1. Getting Cluster id with Tadpole

Function to get the cluster id for clusters

```{r}
# get the cluster id for clusters

# function to plot and get dataframe for cluster_id for Peptide_id
peptide_to_clusters_tadpole <- function(df_filt8,depth="SRF",num_cluster=4L){
  df_SRF8 <- df_filt8 %>% filter(Depth==depth) %>% select(Peptide_id,Time,Ave_abundance_ratio)
  # reshape the dataframe with timepoints
  library(maditr)
  v_names <- c("T1","T2","T3","T4","T5","T6","T7","T8")
  df_SRF8$Time <- factor(df_SRF8$Time, levels=v_names)
  dc_SRF8 <- dcast(df_SRF8, Peptide_id ~ Time, value.var = "Ave_abundance_ratio")
  dc_SRF8_save <- dc_SRF8
  # make Peptide_id more simpler (Pep_1 -> P.1)
  dc_SRF8$Peptide_id <- str_replace(dc_SRF8$Peptide_id,"Pep_","P.")
  # make first column into row name
  dc_SRF8 <- tibble::column_to_rownames(dc_SRF8, var="Peptide_id")
  dim(dc_SRF8)
  # make it into list format
  # l_SRF8 <- setNames(split(dc_SRF8, seq(nrow(dc_SRF8))), rownames(dc_SRF8))
  l_SRF8 <- as.list(as.data.frame(t(dc_SRF8)))
  l_SRF8_z <- zscore(l_SRF8)
  # tadpole
  pc_tp4 <- tsclust(l_SRF8_z, k = num_cluster, type = "t",
    seed = 8, trace = TRUE,
    control = tadpole_control(dc = 1.5,
    window.size = 4L))
  plot(pc_tp4)
  plot(pc_tp4, type="centroid")
  df_cluster <- data.frame(Peptide_id = dc_SRF8_save$Peptide_id, Cluster_id = pc_tp4@cluster)
  return(df_cluster)
}


```

Make dataframe for all depths

```{r}
# 4 clusters
tad_SRF4 <- peptide_to_clusters_tadpole(df_filt8,depth="SRF",num_cluster=4L)
write.table(tad_SRF4,file='tad_SRF4.tsv', sep='\t',row.names = F)

tad_DCL4 <- peptide_to_clusters_tadpole(df_filt8,depth="DCL",num_cluster=4L)
write.table(tad_DCL4,file='tad_DCL4.tsv', sep='\t',row.names = F)

tad_BDCL4 <- peptide_to_clusters_tadpole(df_filt8,depth="BDCL",num_cluster=4L)
write.table(tad_BDCL4,file='tad_BDCL4.tsv', sep='\t',row.names = F)


```






