---
title: "taxonomy_assignment_and_analysis"
author: "KiseokUchicago"
date: "2021-03-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

### Taxonomy assignment to peptides and taxonomy profiling analysis

```{r}
# libraries
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(formattable)
```

### 0. Get the rhythmic(periodic) peptide_ids from the rain analysis (from previous script of "peptide_level_visualization.Rmd")
```{r}
# read in combined table of 24 samples (Replicate B)
df_pep <- read.table("Replicate_ABC67_combined.tsv",sep='\t',header = T)
# Let's try to plot all of them at once
df_all <- df_pep %>% select(Peptide_id, SampleID, Time, Depth) %>% unique() %>% arrange(Peptide_id)
# How many time points did it appear
(df_all2 <- df_all %>% group_by(Peptide_id, Depth) %>% summarize(Timepoints = n_distinct(Time)) %>% ungroup())
df_all2$Depth <- factor(df_all2$Depth, levels = c('SRF','DCL','BDCL'))
# (1) Get list of peptides that are present in 7 or 8 timepoints
df_all2_filt <- df_all2 %>% filter(Timepoints > 6) %>% unique() 
# (2) Let's average the abundance_ratio for each unique peptide
df_int <- df_pep %>% select(Peptide_id, Spec_id, cent.log2.16O.18O.Ratio, SampleID, Time, Depth, Replicate) %>%
  unique() %>% arrange(Peptide_id)
# first average the spectrum_id per sample 
# important to make sure to take the mean of the log raio (instead of taking log2 after averaging the original value(2^x))
df_ave <- df_int %>% group_by(SampleID, Time, Depth, Replicate, Peptide_id) %>%
  summarize(Mean_abundance_ratio=mean(cent.log2.16O.18O.Ratio),
            SD_pep=sd(cent.log2.16O.18O.Ratio)) %>% ungroup()
# Now, average and error by replicate
# SE calculated with Jake Waldbauer's method: sqrt(se(Mean_abundance_ratio)^2 +(mean(SD_pep))^2))
# standard error
se <- function(x) sd(x)/sqrt(length(x))
df_ave_rep <- df_ave %>% group_by(Time, Depth, Peptide_id) %>%
  summarize(Ave_abundance_ratio=mean(Mean_abundance_ratio),
            SE_abundance_ratio=sqrt((se(Mean_abundance_ratio))^2 +(mean(SD_pep))^2)) %>% ungroup()
# why is there no sd for Pep_1004?
df_ave %>% group_by(SampleID) %>% filter(Peptide_id=='Pep_1003') # there is only 1 sample for T1 BDCL 
df_ave_rep2 <- df_ave_rep %>% mutate(error_up=Ave_abundance_ratio+SE_abundance_ratio) %>%
  mutate(error_down=Ave_abundance_ratio-SE_abundance_ratio)
# (3) import time metadata and merge
(df_time <- read.table('data/Time_metadata.txt', header=T, sep='\t'))
df_time$date_time <- str_replace(df_time$date_time, "/2019"," ")
df_time$date_time <- str_replace(df_time$date_time, ":00$","")
# left_join with average data
df_ave2 <- df_ave_rep2 %>% left_join(df_time, by=c('Time'='timepoint'))
head(df_ave2)
# only use peptides that appear 7 or 8 times
df_ave2 %>% select(Peptide_id) %>% unique() %>% dim() # 3233 peptides before
df_filt78 <- inner_join(df_ave2, df_all2_filt, by=c('Peptide_id'='Peptide_id','Depth'='Depth'))
df_filt78$Depth <- factor(df_filt78$Depth, levels = c('SRF','DCL','BDCL'))

# get only 8 ones
df_filt8 <- df_filt78 %>% filter(Timepoints > 7)
df_filt8 %>% select(Peptide_id) %>% unique() %>% dim() # 661 peptides for peptides in all 8 timepoints
df_filt8$Depth <- factor(df_filt8$Depth, levels = c('SRF','DCL','BDCL'))
# rain analysis function
rain_analysis <- function(df_filt8,depth="SRF"){
  # setting our data into input format
  df_8 <- df_filt8 %>% filter(Depth==depth) %>% select(Peptide_id,Time,Ave_abundance_ratio)
  # reshape the dataframe with timepoints
  library(maditr)
  v_names <- c("T1","T2","T3","T4","T5","T6","T7","T8")
  df_8$Time <- factor(df_8$Time, levels=v_names)
  dc_8 <- dcast(df_8, Peptide_id ~ Time, value.var = "Ave_abundance_ratio")
  # make Peptide_id more simpler (Pep_1 -> 1)
  dc_8$Peptide_id <- str_replace(dc_8$Peptide_id,"Pep_","P.")
  # make first column into row name
  dc_8 <- tibble::column_to_rownames(dc_8, var="Peptide_id")
  t <- c(1,0,0,1,0,0,1,0,0,1,0,0,0,0,0,1,0,0,0,1,0,0,1,0,0,1) # adjusted for irregular repeats
  # RAIN analysis
  library(rain)
  rainresult <- rain(t(dc_8), period=8, measure.sequence=t, deltat=1, method='independent')
  return(rainresult)
}

rain_SRF8 <- rain_analysis(df_filt8,depth="SRF") %>% mutate(Depth="SRF") %>% tibble::rownames_to_column(var = "Peptide_id")
rain_DCL8 <- rain_analysis(df_filt8,depth="DCL") %>% mutate(Depth="DCL") %>% tibble::rownames_to_column(var = "Peptide_id")
rain_BDCL8 <- rain_analysis(df_filt8,depth="BDCL") %>% mutate(Depth="BDCL") %>% tibble::rownames_to_column(var = "Peptide_id")
dim(rain_SRF8)
dim(rain_DCL8)
dim(rain_BDCL8)
rain_8 <- rbind(rain_SRF8, rain_DCL8) %>% rbind(rain_BDCL8)
dim(rain_8)
tail(rain_8)

rain_sig <- rain_8 %>% filter(pVal <= 0.25)

color_vec <- c('#fc8d62','#66c2a5','#8da0cb')
depth_vec <- c('SRF','DCL','BDCL')

rain_sig$Depth <- factor(rain_sig$Depth, levels=depth_vec)
rain_sig$Peptide_id <- str_replace(rain_sig$Peptide_id,"P.","Pep_")

```

### 1. Input the taxonomy data (created from linux)
Get Final_pep_seq_ORF_rankedlineage.tsv from 05_taxonomy (at /project2/mlcoleman/project/diel_proteomics_kiseok/20210302_diel_proteomics_replicateABC_67samples) \
This is a tab-delimited dataframe.
Peptide_id    Peptide_sequence    Strain   Species   Genus   Family    Order   Class   Phylum    Phylum    Kingdom

The blanks are denoted as "NULL"
```{r}
# read in the dataframe and change "NULL" to NA
df_peptax <- read.table("data/Final_pep_seq_ORF_rankedlineage.tsv",sep='\t',header = F,na.strings=c("NULL","NA"))

colnames(df_peptax) <- c("Peptide_id","Peptide_sequence","ORF","Strain","Species","Genus","Family","Order","Class","Phylum","Kingdom","Domain")

df_peptax %>% select(Kingdom) %>% unique()
df_peptax %>% select(Domain) %>% unique()

# Are there NAs in the dataframe?
head(df_peptax)
dim(df_peptax)
any(is.na(df_peptax))

# first test with the Surface periodic peptides
rhythm_SRF <- rain_sig %>% filter(Depth=="SRF") %>% select(Peptide_id)
rhy_tax_SRF <- rhythm_SRF %>% left_join(df_peptax, by=c("Peptide_id"="Peptide_id"))

rhy_tax_SRF %>% select(Peptide_id) %>% unique() %>% dim() # 69 peptides

# See the distribution of number of matched ORFs
rhy_t <- rhy_tax_SRF %>% group_by(Peptide_id) %>% summarize(freq = n())

ggplot(rhy_t, aes(freq)) + geom_bar() +
  ylab("Frequency \n") +
  xlab("\n Number of matched ORFs per peptide") +
  # scale_x_continuous(breaks=seq(1,8,1))+
  ggtitle("Histogram of matched ORFs per peptide (Surface) \n")+
  ## adjust positions
  theme(plot.title = element_text(size = 20,hjust = 0.5, family="serif")) + 
  theme(axis.title.x = element_text(size = 15,hjust = 0.5, family="serif")) + 
  theme(axis.title.y = element_text(size = 15,hjust = 0.5, family="serif")) + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust=0.3,size=13, family="serif"))+
  theme(axis.text.y = element_text(size=10))

rhy_tax_SRF %>% left_join(rhy_t, by=c("Peptide_id"="Peptide_id")) %>% filter(freq==1)
rhy_tax_SRF %>% left_join(rhy_t, by=c("Peptide_id"="Peptide_id")) %>% filter(freq==1) %>% dim()
rhy_tax_SRF1 <- rhy_tax_SRF %>% left_join(rhy_t, by=c("Peptide_id"="Peptide_id"))
```
43 out of 69 peptides have no matches to the ORF.\
Let's work with only the peptides that have ORF matches

### 1. Lowest common ancester method assignment


```{r}
# Working only with the peptides that have ORF matches
rhy_tax_SRF2 <- rhy_tax_SRF1 %>% filter(!is.na(ORF))
rhy_tax_SRF2 %>% select(freq) %>% unique() %>% arrange(freq)
# NA in taxonomy?
rhy_tax_SRF2 %>% filter(is.na(Strain))

# plot the scientific name level histogram
df.sci <- rhy_tax_SRF2 %>% count(Peptide_id, Strain)

# we need to group by Peptides
df.sci.rel <- df.sci %>%  
  group_by(Peptide_id) %>%                               
  mutate(RelAbundance = n*100/sum(n))  # Transform to rel. abundance

# df.sci.rel[df.sci.rel$RelAbundance < 5,]$Phylum <- 'Low abundance'

### let's test
ord <- df.sci.rel %>% group_by(Strain) %>% summarise(sumRel = sum(RelAbundance)) %>% arrange(sumRel)
vec <- ord$Strain
rev(tail(vec,10))
head(vec)

df.sci.rel$Strain <- factor(df.sci.rel$Strain, levels = vec) 

my_color <- c(
  "black","#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange")

my_color2 <- c(
  "#CBD588", "#DA5724","#AD6F3B", "#673770","#D14285", 
  "#652926", "#C84248", "#8569D5", "#5E738F","#D1A33D", 
  "#8A7C64", "#599861", "#616163",  "#FFCDB2", "#242F40", 
  "#6D9F71", "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#9FB1BC', 'springgreen', '#E2C044', '#5BC0EB', 'pink', 
  "orange", "#CBD588", "#5F7FC7",  
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "#5F7FC7", "orange",  
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861",
  "#616163",  "#FFCDB2", "#242F40", "#6D9F71", 
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724",
  "#CBD588", "orange")

# plot relative abundance
df.sci.rel.p1 <- ggplot(df.sci.rel, aes(x=Peptide_id, y = RelAbundance, fill = Strain)) + 
  geom_bar(stat="identity", width = 0.8, position = 'stack', colour="black") +
  #scale_fill_discrete() +
  scale_fill_manual(values = my_color2) +
  
  xlab('')+
  ylab("Relative Abundance (%) \n") +
  ggtitle("Matched taxa for each peptide \n") +
  ## adjust positions
  guides(fill = guide_legend(ncol = 7,reverse = T))+
  theme(legend.position="bottom") +
  theme(plot.title = element_text(size = 20,hjust = 0.5, face='bold')) + 
  theme(axis.title.x = element_text(size = 15,hjust = 0.5, face='bold')) + 
  theme(axis.title.y = element_text(size = 13,hjust = 0.5, face='bold')) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0.95, vjust=0.1,size=12, face='bold',color='black'))+
  theme(axis.text.y = element_text(size=15, face='bold',color='black'))+
  scale_y_continuous(breaks=seq(0,1000,100))+
  theme(panel.grid.major = element_blank()) +
  theme(panel.grid.minor = element_blank(), panel.background=element_blank(),panel.border=element_blank(), plot.background=element_blank())

df.sci.rel.p1


```

Let's go to the upper taxa level and see. \

Making of function


```{r}
# function for plotting

rhythmic_tax_profile <- function(rain_sig, depth="SRF", df_peptax, tax_level,legend_coln=7){
  # tax_level=c("Strain","Species","Genus","Family","Order","Class","Phylum","Kingdom","Domain")
  tax_col=tax_level
  # first test with the Surface periodic peptides
  rhythm_SRF <- rain_sig %>% filter(Depth==depth) %>% select(Peptide_id)
  rhy_tax_SRF <- rhythm_SRF %>% left_join(df_peptax, by=c("Peptide_id"="Peptide_id"))
  rhy_tax_SRF %>% select(Peptide_id) %>% unique() %>% dim() # 69 peptides
  # Working only with the peptides that have ORF matches
  rhy_tax_SRF2 <- rhy_tax_SRF %>% filter(!is.na(ORF))
  # NA in taxonomy?
  rhy_tax_SRF2 %>% filter(is.na(!!sym(tax_col)))

  # plot the scientific name level histogram
  df.sci <- rhy_tax_SRF2 %>% count(Peptide_id, !!sym(tax_col))

  # we need to group by Peptides
  df.sci.rel <- df.sci %>%  
    group_by(Peptide_id) %>%                               
    mutate(RelAbundance = n*100/sum(n))  # Transform to rel. abundance
  
  ### let's test
  ord <- df.sci.rel %>% group_by(!!sym(tax_col)) %>% summarise(sumRel = sum(RelAbundance)) %>% arrange(sumRel)
  vec <- ord[,1]
  tail(vec)
  
  # plot relative abundance
  df.sci.rel.p1 <- ggplot(df.sci.rel, aes_string(x="Peptide_id", y = "RelAbundance", fill = tax_col)) + 
    geom_bar(stat="identity", width = 0.8, position = 'stack', colour="black") +
    scale_fill_manual(values = my_color2) +
    xlab('')+
    ylab("Relative Abundance (%) \n") +
    ggtitle(paste0("Matched taxa for rhythmic peptides (Depth: ",depth," / Level: ",tax_col,")")) +
    ## adjust positions
    guides(fill = guide_legend(ncol = legend_coln,reverse = T))+
    theme(legend.position="bottom") +
    theme(plot.title = element_text(size = 20,hjust = 0.5, face='bold')) + 
    theme(axis.title.x = element_text(size = 15,hjust = 0.5, face='bold')) + 
    theme(axis.title.y = element_text(size = 13,hjust = 0.5, face='bold')) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 0.95, vjust=0.4,size=12, face='bold',color='black'))+
    theme(axis.text.y = element_text(size=15, face='bold',color='black'))+
    scale_y_continuous(breaks=seq(0,100,10))+
    theme(panel.grid.major = element_blank()) +
    theme(panel.grid.minor = element_blank(), panel.background=element_blank(),panel.border=element_blank(), plot.background=element_blank())
  
  return(df.sci.rel.p1)
}
```

Plotting for different levels
```{r}
# SRF rhythmic taxa - Strain level
rhythmic_tax_profile(rain_sig, depth="SRF", df_peptax, tax_level="Strain", legend_coln=7)

# SRF rhythmic taxa - Species level
rhythmic_tax_profile(rain_sig, depth="SRF", df_peptax, tax_level="Species", legend_coln=7)

# SRF rhythmic taxa - Genus level
rhythmic_tax_profile(rain_sig, depth="SRF", df_peptax, tax_level="Genus", legend_coln=7)

# SRF rhythmic taxa - Family level
rhythmic_tax_profile(rain_sig, depth="SRF", df_peptax, tax_level="Family", legend_coln=7)

# SRF rhythmic taxa - Order level
rhythmic_tax_profile(rain_sig, depth="SRF", df_peptax, tax_level="Order", legend_coln=7)

# SRF rhythmic taxa - Phylum level
rhythmic_tax_profile(rain_sig, depth="SRF", df_peptax, tax_level="Phylum", legend_coln=7)

# SRF rhythmic taxa - Kingdom level
rhythmic_tax_profile(rain_sig, depth="SRF", df_peptax, tax_level="Kingdom", legend_coln=7)

# SRF rhythmic taxa - Domain level
rhythmic_tax_profile(rain_sig, depth="SRF", df_peptax, tax_level="Domain", legend_coln=7)


```



